### 2021-02-09 final_report/include/CE.tex
quickly reviewed chapters about CP and CE

\section{Method and implementation} \label{ce}
For the reasons discussed in Section \ref{ourapproach}, our syntax-based approach to CE is based on comparing, instead on ASTs, dependency trees, and UD trees in particular.
The task of dealing with UD trees is facilitated by the existence of a series of Haskell modules written in the context of the development of the aforementioned \texttt{gf-ud} and of some preliminary unpublished CA experiments\footnote{Aarne Ranta and Prasanth Kolachina, code available at \url{https://github.com/aarneranta/concept-alignment/tree/master/old_ca_versions/v1}}. \smallskip

>>>>

\section{Method and implementation} \label{ce}
For the reasons discussed in Section \ref{ourapproach}, our syntax-based approach to CE is based on comparing, instead on ASTs, dependency trees, and UD trees in particular.
The task of dealing with UD trees is facilitated by the existence of a series of Haskell modules written in the context of the development of the aforementioned \texttt{gf-ud} and of some preliminary unpublished CA experiments\footnote{Aarne Ranta and Prasanth Kolachina, code available at \url{github.com/aarneranta/concept-alignment/tree/master/old_ca_versions/v1}}. \smallskip

### 2021-02-09 final_report/include/CP.tex
quickly reviewed chapters about CP and CE

% CREATED BY DAVID FRISK, 2016
\chapter{Concept Propagation} \label{ch4}
After a set of concepts is known, it can be useful to be able to look for the corresponding concrete expressions in a new language. In this chapter, we describe our approach to this task and present an evaluation of the corresponding module.

@@

\paragraph{Propagation with texts in different domains}
When it comes to the first set of experiments, where we use PUD data, being the subsets of selected sentences extremely small and having the correctness of most alignments been already assessed in previous experiments, it is easy to work with all possible pairs of languages. The results, summarized in Table \ref{tcp2pud}, show that, in general, only a small fraction of the extracted alignments are actually propagated in this case. As mentioned above, this is to be expected since different sentences belong to different domain. A closer look to the set of alignments returned by propagation, the vast majority of which has single words as members, shows that most of them are common function words, which, alone, are not always relevant for a GF-based MT pipeline. 
Common determiners such as articles, for instance, can be better handled by RGL grammars alone. 
Prepositions, on the other hand, are only useful with some information about the context in which they occur in, since it is customary for them to be used differently across different, even closely related languages. 
However, some common content words, like $\langle$\textit{always, sempre, alltid}$\rangle$, $\langle$\textit{new, nuovi, nya}$\rangle$ and $\langle$\textit{same, stesso, samma}$\rangle$ are also aligned correctly.

@@

    More than the quantitative results, however, it is interesting to take a look at some of the propagated concepts. Because the text are similar both stylistically and, most importantly, in terms of content, the program is in this case also able to find trilingual correspondences between content words, sometimed specific of the domain. Examples of such alignments are $\langle$\textit{skills, capacità, färdigheter}$\rangle$, $\langle$\textit{functional, funzionale, funktionell}$\rangle$, $\langle$\textit{exam, prova, tentamen}$\rangle$, $\langle$\textit{course, corso, kurs}$\rangle$, $\langle$\textit{prerequisites, prerequisiti, behörigheter}$\rangle$, $\langle$\textit{lectures, lezioni, föreläsningar}$\rangle$ and $\langle$\textit{knowledge, conoscenza, kunskap}$\rangle$. \smallskip

    Sometimes, however, the meaning of a word is highly context-dependent and as a consequence, even if both the pairs of extracted and propagated alignments are correct, the corresponding trilingual alignment we could infer isn't (e.g. $\langle$\textit{learning, conoscere}$\rangle$ and $\langle$\textit{learning, inlärning}$\rangle$). \smallskip

    Finally, in some rare cases, even correspondences regarding longer expressions, such as $\langle$\textit{the aim of the course, l'obiettivo del corso, syftet med kursen}$\rangle$, can be found. In this case, the English-Italian correspondence is not especially useful, since the phrase could be translated almost word by word between the two languages, but the translation to Swedish is not literal, thus making the alignment useful for the purposes of MT.

>>>>

% CREATED BY DAVID FRISK, 2016
\chapter{Concept Propagation} \label{ch4}
After a set of concepts is known, it can be useful to look for the corresponding concrete expressions in a new language. In this chapter, we describe our approach to this task and present an evaluation of the corresponding module.

@@

\paragraph{Propagation with texts in different domains}
When it comes to the first set of experiments, where we use PUD data, being the subsets of selected sentences extremely small and having the correctness of most alignments been already assessed in previous experiments, it is easy to work with all possible pairs of languages. The results, summarized in Table \ref{tcp2pud}, show that, in general, only a small fraction of the extracted alignments are actually propagated in this case. As mentioned above, this is to be expected since different sentences belong to different domain. A closer look to the set of alignments returned by propagation, the vast majority of which has single words as members, shows that most of them are common function words, which, alone, are not always relevant for a GF-based MT pipeline. 
Common determiners such as articles, for instance, can be better handled by RGL grammars alone. 
Prepositions, on the other hand, are only useful with some information about the context in which they occur in, since it is customary for them to be used differently across different, even closely related languages. 
However, some common content words, like $\langle$\textit{``always'', ``sempre'', ``alltid''}$\rangle$, $\langle$\textit{``new'', ``nuovi'', ``nya''}$\rangle$ and $\langle$\textit{``same'', ``stesso'', ``samma''}$\rangle$ are also aligned correctly.

@@

    More than the quantitative results, however, it is interesting to take a look at some of the propagated concepts. Because the text are similar both stylistically and, most importantly, in terms of content, the program is in this case also able to find trilingual correspondences between content words, sometimed specific of the domain.

    Examples of such alignments are $\langle$\textit{``skills'', ``capacità'', ``färdigheter''}$\rangle$, $\langle$\textit{``functional'', ``funzionale'', ``funktionell''}$\rangle$, $\langle$\textit{``exam'', ``prova'', ``tentamen''}$\rangle$, $\langle$\textit{``course'', ``corso'', ``kurs''}$\rangle$, $\langle$\textit{``lectures'', ``lezioni'', ``föreläsningar''}$\rangle$, $\langle$\textit{``prerequisites'', ``prerequisiti'', ``behörigheter''}$\rangle$ and $\langle$\textit{``knowledge'', ``conoscenza'', ``kunskap''}$\rangle$. \smallskip

    Sometimes, however, the meaning of a word is highly context-dependent and as a consequence, even if both the pairs of extracted and propagated alignments are correct, the corresponding trilingual alignment we could infer isn't (e.g. $\langle$\textit{``learning'', ``conoscere''}$\rangle$ and $\langle$\textit{``learning'', ``inlärning''}$\rangle$). \smallskip

    Finally, in some rare cases, even correspondences regarding longer expressions, such as $\langle$\textit{``the aim of the course'', ``l'obiettivo del corso'', ``syftet med kursen''}$\rangle$, can be found. In this case, the English-Italian correspondence is not especially useful, since the phrase could be translated almost word by word between the two languages, but the translation to Swedish is not literal, thus making the alignment useful for the purposes of MT.
